##Reproducible Research: Project Assignment 1

####Loading required Libraries

```{r loadlib, echo=FALSE, results="hide"}
library(knitr)
library(dplyr)
library(xtable)
```
```{r setoptions, echo=FALSE}
opts_chunk$set(echo=TRUE, results="show")
```

####Load the data from activity.csv file. Printing summary of data.

```{r load}
all_data <- read.csv("activity.csv",header = TRUE,stringsAsFactors = FALSE)
str(all_data)
```


####Function to draw histogram by date

```{r fun_hist}
drawhist <- function(data) {
  by_date <- group_by(data,date)
  sum_data <- summarize(by_date,"Sum" = sum(steps),"Mean" = mean(steps), 
                       "Median" = median(steps) )
  xt <- xtable(sum_data)
  print(xt,type="html")
  hist(sum_data$Sum,main="Steps Taken Per Day",xlab="Number of Steps",ylab="Days")
}
```

####Function to draw line plot by 5-min interval.

```{r fun_line}
drawLine <- function(data,title) {
  act_by_intv <- group_by(data,interval)
  step_intv <- summarize(act_by_intv,"steps" = mean(steps,na.rm = TRUE))
  plot(step_intv$interval,step_intv$steps,xlab="Interval",ylab="Average Steps",
     main=title, type="l")
  max_intv = step_intv[step_intv$steps==max(step_intv$steps),]$interval
  max_intv
  
}
```

####Remove NA's and create good data.

```{r remNA} 
  good<-complete.cases(all_data)
  gdata <- all_data[good,]
```



####Histogram of Total steps taken per day ignoring NA's.
####Mean and Median shown for steps taken each day.

```{r good_data_anly, fig.width=12,fig.height=8,results="asis"}
  drawhist(gdata)
```


####Average number of steps taken (averaged across all days) versus the 5-minute intervals

```{r timeplot,,fig.width=12,fig.height=8}
  max_intv=drawLine(gdata,"Average Steps in 5min Interval")
```

**The 5-minute interval on an average across all days** 
**taking maximun number of steps is : `r max_intv`**


**The missing values in dataset are `r nrow(all_data) - nrow(gdata)`**


####Filling the NA's with mean for that date and creating new data set.
####The mean of steps is taken from good data from each day then for each
####data element which has NA same is used.


```{r fill_NA}
fdata <- all_data
by_date <- group_by(gdata,date)
avg_data <- summarize(by_date,"avg" = mean(steps))
fdata[!good,]$steps <- avg_data[fdata[!good,]$date,]$avg
```

###Histogram for average steps per day using filled data

```{r hist_with_fil, fig.width=12,fig.height=8,results="asis"}
drawhist(fdata)
```

**It is observed that there is no impact in mean values of the data without NA's**
**versus the filled up data.**
**There is no impact of missing data on estimates of total daily number of steps.**



####Extact Weekend and Weekday data 
####Time series plot for average number of steps taken per 5-minute interval 
####across weekdays and weekends 
 

```{r weekplots, fig.width=12, fig.height=12}

fact <- weekdays(as.Date(fdata$date))
fact[fact=="Saturday" | fact=="Sunday"] <- "Weekend"
fact[fact !="Weekend"] <- "Weekday"
wdata <- cbind(fdata,fact);


par(mfrow=c(2,1))
drawLine(wdata[wdata$fact=="Weekend",],"Avg Steps in Weekend")
drawLine(wdata[wdata$fact=="Weekday",],"Avg Steps in Weekday")


```