---
title: "PracticalMachineLearning-CourseAssignment"
author: "Rakesh Bhatnagar"
date: "10 July 2016"
output: html_document
---

Load the Training & Testing Data. The data is pre-downloaded from below websites.
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv and 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv websites.

```{r}
library(caret)
setwd("D:/Rakesh/MYWORK/RSpace/PML")
traindata <- read.csv("pml-training.csv",stringsAsFactors=FALSE,na.strings=c("","NA"))

testingdata <- read.csv("pml-testing.csv",stringsAsFactors=FALSE,na.strings=c("","NA"))

```

Perform data preprocessing. 
Remove the columns that are not required for building model.

```{r}
traindata <- traindata[,-c(1:7)]
testingdata <- testingdata[,-c(1:7)]
```

Few columns have good amount of missing data. These can be removed for model building
Removing the columns that have 95% or more missing data.

```{r}
str(traindata)
nr<-dim(traindata)[1]
traindata <- traindata[ , colSums(is.na(traindata)) <= 0.95*nr]
preObj <- preProcess(traindata[,-53],method=c("center","scale"))
trainScaleData<-predict(preObj,traindata[,-53])
testingdata<-testingdata[,colnames(trainScaleData)]
alltrain <- data.frame(trainScaleData,classe=traindata[,53])
```

Build the validation data from the training data. Taking 50% of the training data to be validation.

```{r}
indxTrain<-createDataPartition(y=alltrain$classe,p=0.5,list=FALSE)
training<-traindata[indxTrain,]
validation<-traindata[-indxTrain,]
str(training)
```

Since the problem is to predict the class. We can start by fitting the basic tree model.

```{r}
mod1 <- train(classe~.,"rpart",data=training)
print(mod1$finalModel)
prediction <- predict(mod1,validation)
confusionMatrix(prediction,validation$classe)
plot(mod1$finalModel,uniform=TRUE,main="Classification Tree")
text(mod1$finalModel,use.n = TRUE, all=TRUE, cex=.8)
```
As can be seen by the above output the accuracy of the model is low.
Let is now try to fit the random forest model.

```{r}
mod2 <- train(classe~.,"rf",data=training)
prediction2 <- predict(mod2,validation)
confusionMatrix(prediction2,validation$classe)
```

The model generated through random forest is very accurate but slow in generation.
Now we can apply this model on the test data.

```{r}
prediction3 <- predict(mod2,testingdata)
prediction3
```

The output gives the prediction classes for the 20 observations from testing data set.
